---
title: "Pract2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r, include=FALSE}
# Apply the EnviroStat package version of the early WLS deformation
# algorithm
library(EnviroStat)
library(MASS)

source("http://www.stat.washington.edu/peter/PASI/deformloglik.l1.R")
source("http://www.stat.washington.edu/peter/591/labs/Lab2/visualize.tps.warps.R")
source("http://www.stat.washington.edu/peter/591/labs/Lab2/draw2.R")  # to replace a function in EnviroStat
######################################################
# For reading the French precipitation dataset
plot.title <- "Pluie.LanguedocRoussillon"
Pluie <- read.csv("http://www.stat.washington.edu/peter/PASI/pluieNA.novdec.csv")
Pluie <- Pluie[, -1]
PluieStations <- read.csv("http://www.stat.washington.edu/peter/PASI/pluieStations.Practicum.csv")
X <- as.matrix(PluieStations[, c("x", "y")])
dimnames(X) <- list(PluieStations$numero, c("x", "y"))
# Feel free to look at the other columns of PluieStations to see, for
# example, site names.  Take a quick look at the configuration of the
# monitoring sites
plot(X[, 1], X[, 2], asp = 1, type = "n")
text(X[, 1], X[, 2], dimnames(X)[[1]])
S <- cov(log(Pluie + 1), use = "pair")  # Pairwise covariances for log precip
t <- nrow(Pluie)
#######################################################
Xs <- scale(X,scale=FALSE)  
Xs <- Xs/sqrt(sum(Xs^2))
Xsvd <- svd(Xs)
# Here we make sure diaonal elements of rotation matrix are positive
# by scaling columns by +/- 1.  This is to keep the rotation from flipping
# the orientation entirely.
Xsvd$v[,1] <- Xsvd$v[,1] * sign(Xsvd$v[1,1]) # 
Xsvd$v[,2] <- Xsvd$v[,2] * sign(Xsvd$v[2,2])
Xt <- Xs %*% Xsvd$v
xscaling <- list(mean=attr(Xs,"scaled:center"),
                 #scale=attr(Xs,"scaled:scale")*sqrt(2*(n-1)),
                 scale=1,
                 rotation=Xsvd$v)
X0 <- X   # Save initial coordinates
X <- Xt   # Replace X with center, scaled, rotated coordinates
n <- nrow(X)
# Take a quick look at the standardized configuration of the monitoring sites
# for comparison to what you looked at above.
plot(X[,1],X[,2],asp=1,type="n")
text(X[,1],X[,2],dimnames(X)[[1]])
################################################
# Part 1 #######################################
################################################
Cor <- S/sqrt(diag(S) %o% diag(S))
Disp <- 2 - 2 * Cor
Gdist <- as.matrix(dist(X, upper = T, diag = T))
plot(Gdist, Disp)
################################################
h.lt <- Gdist[row(Gdist) < col(Gdist)]
disp.lt <- Disp[row(Disp) < col(Disp)]
variogfit <- EnviroStat::Fvariogfit3(disp.lt, h.lt, a0 = 0.1, t0 = 1, verbose = T)
x <- seq(min(h.lt), max(h.lt), 0.01)
a0 <- variogfit$a[1]
t0 <- variogfit$t0

plot(-0.2, 0, xlim = c(0, max(h.lt)), ylim = c(0, 2), xlab = "Dist", ylab = "Dispersion", 
     type = "n", main = c("Intersite dispersion vs intersite distance"))
for (i in 1:(n - 1)) for (j in (i + 1):n) points(Gdist[i, j], Disp[i, j])
lines(x, a0 + (2 - a0) * (1 - exp(-(t0 * x))), col = "red")
# Step 1
source("http://www.stat.washington.edu/peter/591/labs/Lab2/plotCoordChange.R")
source("http://www.stat.washington.edu/peter/591/labs/Lab2/plotDispersionDist.R")
assignInNamespace("plotCoordChange", plotCoordChange, ns = "EnviroStat")
assignInNamespace("plotDispersionDist", plotDispersionDist, ns = "EnviroStat")
sg.est <- Falternate3(Disp, X, max.iter = 100, alter.lim = 25, model = 1, lambda = 0)
######################################
source("http://www.stat.washington.edu/peter/591/labs/Lab2/Ftransdraw.R")
assignInNamespace("Ftransdraw", Ftransdraw, ns = "EnviroStat")
coords.grid <- Fmgrid(range(X[, 1]), range(X[, 2]))
par(mfrow = c(1, 2))
# temp <- setplot(X, axes = TRUE)
plot(X, type = "n", xlab = "", ylab = "", asp = T)
deform <- Ftransdraw(disp = Disp, Gcrds = X, MDScrds = sg.est$ncoords, gridstr = coords.grid)
# 0.001
penParam=40
lam <- penParam
sg.est <- Falternate3(Disp, X, max.iter = 100, alter.lim = 25, model = 1, lambda = lam)
Tspline <- sinterp(X, sg.est$ncoords, lam = 0)
temp <- Ftransdraw(disp = Disp, Gcrds = X, MDScrds = Tspline$y, gridstr = coords.grid, 
                   eye = F)
########################################
par(mfrow = c(1, 1), mgp = c(2, 0.75, 0))
Tgrid <- bgrid(start = c(0, 0), xmat = X, coef = Tspline$sol)
```

## Question 1

```{r}
plot(X, xlab = "", ylab = "", type = "n", asp = TRUE)
text(X, labels = 1:nrow(X), cex = 0.75)
Bgrid.title <- paste("WLS Bgrid: ", plot.title, "\n lambda=", lam, sep = "")
drawout <- draw2(Tgrid, fs = TRUE, lwidth = c(2, 2), lcolor = topo.colors(5), 
                 legend = TRUE)
title(main = Bgrid.title)
title(sub = "Lines are colored by directional magnitude of the
      principal axes of the deformation")
```

#In what area(s) of the map is spatial correlation strongest and what area(s) is it weakest?
Spatial correlation is strongest nearer the coast and weaker further away from the coast, particularly in the North-West

#Where is local anisotropy greatest?
Along the coast line

#Are there regions where the local correlation structure appears nearly isotropic?

#Can you relate any of this estimated structure to the topography in the tourist map shown above?
